### 2019-11-06 thesis/txt/2-models.tex
adress some linter warnings from lacheck and style-check.rb

\chapter{Chunk Retrieval Techniques for Build Logs}
\plan{whole chapter: some reordering and reviewing to do, write about 3 pages about ir technique model classes and ir tasks and other techniques}
\plan{write, 1/3 page, blocked by: rest of chapter}
\label{sec:ir-techniques}
%\includegraphics[page=1, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/RetreatSeptemberSlides.pdf}
Many different information retrieval techniques ... this section describes the conceptional / theoretical results of our thesis ...
We create a model to characterize BLIE techniques so we can discuss their similarities and differences more clearly.
Additionally, we present a complementary model for IE tasks, to describe use cases for BLIE techniques in a structured way.
From our analysis of build logs, we share our notion of retrievable information within a build log.

@@

\begin{itemize}
	\item Pulling the new, changed version of the source code onto the server.
	\item Running static analysis tools~\cite{zampetti2017open}.
	\item Building the software, i.e. compiling and packaging it~\cite{phillips2014understanding}.
	\item Running automated tests~\cite{beller2017oops}.
	\item Deployment of the build artifact~\cite{schermann2016empirical}.
\end{itemize}

@@

A BLIRT
\begin{itemize}
	\item targets a specific \textbf{BuildLogInformation}.
	\item identifies the text part to retrieve through an \textbf{IdentificationTechnique}.
	\item is configured by a \textbf{Configuration}
	      \begin{itemize}
		      \item each configuration addresses a specific \textbf{Scope}.
		            The Scope can be a specific project, a tool involved in the build, a programming language or global, configuring a retrieval technique for all possible build logs.
	      \end{itemize}
	\item can retrieve text passages with a specific \textbf{Granularity}, i.e. the smallest extractable text piece.
	      The Granularity can be at the level of characters, lines or paragraphs.
\end{itemize}

@@

\emph{Programming by Example (PBE)} is a technique where programs are synthesized according to in and output examples provided by the user.
It enables users to create programs without requiring programming knowledge~\cite{mayer2015user}.
Applied to the task of text extraction through regular expressions it relieves the developer from having to understand the whole document structure to solve a single extraction task~\cite{le2014flashextract:}.
We chose to investigate the applicability of generating regular expression programs by example to extracting information from build logs.
In this work, we refer to our interpretation of this technique as \emph{PBE}.
In the following we will explain the concept of how we apply PBE to information retrieval from CI build logs, the concrete implementation is described in~\ref{sec:impl-pbe}.
Figure~\ref{fig:prose-explanation} shows the instantiation of the BLIRT model for PBE.

\paragraph{Method}
In/Output examples (I/O examples) are the main driver of Programming by Example.
When extracting information from build logs the \emph{input} is always a whole build log, i.e. the whole text of the build log file without any preprocessing.
In the implementation, we had to apply a few text replacements as described in~\ref{sec:impl-preprocessing}.
The \emph{output} is always a substring of the log file text, representing the substring that should be extracted by the synthesized program when given the corresponding input file.
One or multiple I/O examples are needed as configuration for PBE, they are used to define substring of a log should be extracted.
Our approach is based on the Flash Extract DSL, which in turn is based on the FlashMeta algorithm.
Both are described in Section~\ref{sec:rw-prose}.
%For execution PBE takes a build file as input and returns a substring of the build file content.

@@

\paragraph{Method}
To configure a retrieval though text similarity we chose to use the same concept of I/O examples as for PBE.
The lines of the output strings of a given example set define our search query.

@@

\paragraph{Log Diff}
Amar et al. used a technique based on line diffs to retrieve relevant lines from a failed build log~\cite{amar2019mining}, as we describe in more detail in Section~\ref{sec:rw-bl-analysis}.
Figure~\ref{fig:diff-technique-model} shows how this technique would be in our BLIRT model.
\paragraph{Random Line Retrieval (RLR)}
In our evaluation we want to compare against a basic deadline of randomly extracted lines.
The random line retrieval technique (RLR) averages the number of lines in the outputs of the configuring examples and extracts this many lines randomly from the build log.

>>>>

\chapter{Chunk Retrieval Techniques for Build Logs}
\plan{whole chapter: some reordering and reviewing to do, write about 3 pages about ir technique model classes and ir tasks and other techniques}
\plan{write, 1/3 page, blocked by: rest of chapter}
\label{sec:ir-techniques}
%\includegraphics[page=1, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/RetreatSeptemberSlides.pdf}
Many different information retrieval techniques\textellipsis this section describes the conceptional / theoretical results of our thesis\textellipsis
We create a model to characterize BLIE techniques so we can discuss their similarities and differences more clearly.
Additionally, we present a complementary model for IE tasks, to describe use cases for BLIE techniques in a structured way.
From our analysis of build logs, we share our notion of retrievable information within a build log.

@@

\begin{itemize}
	\item Pulling the new, changed version of the source code onto the server.
	\item Running static analysis tools~\cite{zampetti2017open}.
	\item Building the software, i.e.\ compiling and packaging it~\cite{phillips2014understanding}.
	\item Running automated tests~\cite{beller2017oops}.
	\item Deployment of the build artifact~\cite{schermann2016empirical}.
\end{itemize}

@@

A BLIRT
\begin{itemize}
	\item targets a specific \textbf{BuildLogInformation}.
	\item identifies the text part to retrieve through an \textbf{IdentificationTechnique}.
	\item is configured by a \textbf{Configuration}
	      \begin{itemize}
		      \item each configuration addresses a specific \textbf{Scope}.
		            The Scope can be a specific project, a tool involved in the build, a programming language or global, configuring a retrieval technique for all possible build logs.
	      \end{itemize}
	\item can retrieve text passages with a specific \textbf{Granularity}, i.e.\ the smallest extractable text piece.
	      The Granularity can be at the level of characters, lines or paragraphs.
\end{itemize}

@@

\emph{Programming by Example (PBE)} is a technique where programs are synthesized according to in and output examples provided by the user.
It enables users to create programs without requiring programming knowledge~\cite{mayer2015user}.
Applied to the task of text extraction through regular expressions it relieves the developer from having to understand the whole document structure to solve a single extraction task~\cite{le2014flashextract:}.
We chose to investigate the applicability of generating regular expression programs by example to extracting information from build logs.
In this work, we refer to our interpretation of this technique as \emph{PBE}.
In the following we will explain the concept of how we apply PBE to information retrieval from CI build logs, the concrete implementation is described in~\ref{sec:impl-pbe}.
Figure~\ref{fig:prose-explanation} shows the instantiation of the BLIRT model for PBE\@.

\paragraph{Method}
In/Output examples (I/O examples) are the main driver of Programming by Example.
When extracting information from build logs the \emph{input} is always a whole build log, i.e.\ the whole text of the build log file without any preprocessing.
In the implementation, we had to apply a few text replacements as described in~\ref{sec:impl-preprocessing}.
The \emph{output} is always a substring of the log file text, representing the substring that should be extracted by the synthesized program when given the corresponding input file.
One or multiple I/O examples are needed as configuration for PBE, they are used to define substring of a log should be extracted.
Our approach is based on the Flash Extract DSL, which in turn is based on the FlashMeta algorithm.
Both are described in Section~\ref{sec:rw-prose}.
%For execution PBE takes a build file as input and returns a substring of the build file content.

@@

\paragraph{Method}
To configure a retrieval though text similarity we chose to use the same concept of I/O examples as for PBE\@.
The lines of the output strings of a given example set define our search query.

@@

\paragraph{Log Diff}
Amar et al.\ used a technique based on line diffs to retrieve relevant lines from a failed build log~\cite{amar2019mining}, as we describe in more detail in Section~\ref{sec:rw-bl-analysis}.
Figure~\ref{fig:diff-technique-model} shows how this technique would be in our BLIRT model.
\paragraph{Random Line Retrieval (RLR)}
In our evaluation we want to compare against a basic deadline of randomly extracted lines.
The random line retrieval technique (RLR) averages the number of lines in the outputs of the configuring examples and extracts this many lines randomly from the build log.

### 2019-11-06 thesis/txt/4-tool.tex
fixes from Taicos introduction review
loads of corrections in ch2, literature, dev mails geschrieben und ausgewertet
incorporate kw 44 structure feedback in thesis, some plot advancements
adress some linter warnings from lacheck and style-check.rb

\chapter{Information Retrieval Tool}
\label{sec:implementation}
We implemented prototypes for our three focus information retrieval techniques to evaluate them on our collected build logs.
This section will present our unified interface and describe the implementation of each of the techniques.

@@

\section{Common Interface}
\begin{lstlisting}
Usage: ruby run-extraction.rb -a analyze -t <technique: ir, pbe, random> -e <example_set> -p <path_to_file_to_analyze>
       ruby run-extraction.rb -a evaluate -t <technique: ir, pbe, keyword, random> -e <example_set> -s <selection_technique> -l <step_count_for_learning> -c <test_count>
       ruby run-extraction.rb -a analyze -t keyword -k <keyword, keyword, ...> -p <path_to_file_to_analyze>
       ruby run-extraction.rb -a analyze -t regex -r <regex to match extraction> -p <path_to_file_to_analyze>

@@

\section{Overall Implementation Remarks}
There are multiple implementation specialities that span over all languages, frameworks and implemented techniques.
The following section discusses preprocessing and escaping of the parsed build log files / example sets, as well as ... \todo{more?}

\subsection{Preprocessing Log Text}
\label{sec:impl-preprocessing}

\section{PROSE Program Synthesis}
\label{sec:impl-pbe}
The first technique we implemented was the PROSE regular expression program synthesis. As it is based on the PROSE library from Microsoft provided in C\# we also chose C\# for our tool. 
implementation details :)

@@

\section{IR Similarity}
impl details,

@@

\section{Keyword search}
impl details,

>>>>

\chapter{Tool Implementation}
\label{sec:implementation}
We implemented prototypes for our three focus information retrieval techniques to evaluate them on our collected build logs.
This section will present our unified interface and describe the implementation of each of the techniques.

@@

%\section{Common Interface}
\begin{lstlisting}
Usage: ruby run-extraction.rb -a analyze -t <technique: ir, pbe, random> -e <example_set> -p <path_to_file_to_analyze>
       ruby run-extraction.rb -a evaluate -t <technique: ir, pbe, keyword, random> -e <example_set> -s <selection_technique> -l <step_count_for_learning> -c <test_count>
       ruby run-extraction.rb -a analyze -t keyword -k <keyword, keyword, ...> -p <path_to_file_to_analyze>
       ruby run-extraction.rb -a analyze -t regex -r <regex to match extraction> -p <path_to_file_to_analyze>

@@

%\section{Overall Implementation Remarks}
There are multiple implementation specialities that span over all languages, frameworks and implemented techniques.
The following section discusses preprocessing and escaping of the parsed build log files / example sets, as well as\textellipsis \todo{more?}

\subsection*{Preprocessing Log Text}
\label{sec:impl-preprocessing}

%\section*{PBE}
\label{sec:impl-pbe}
The first technique we implemented was the PROSE regular expression program synthesis. As it is based on the PROSE library from Microsoft provided in C\# we also chose C\# for our tool. 
implementation details :)

@@

%\section*{IR Similarity}
\label{sec:impl-ts}
impl details,

@@

%\section*{Keyword search}
\label{sec:impl-skws}
impl details,

### 2019-11-06 thesis/txt/5-study.tex
adress some linter warnings from lacheck and style-check.rb

The examples are sorted chronologically, i.e. a technique is configured with examples from the directly preceding build logs.
For each example set and each technique, we select one to five consecutive I/O examples as configuration, the \emph{configuring I/O examples}, and run the chunk retrieval on the following I/O example, the \emph{test I/O example}.

For each of our evaluation runs we measure the desired vs. actual output lines of the chunk retrieval.
The desired output is given by the output of the test I/O example, which acts as the oracle for our evaluation.
From this we obtain true/false positives/negatives and calculate precision, recall and accuracy.
We use these values to answer \textbf{RQ2.2}.
If all desired lines are extracted, i.e. recall is one, we say a chunk retrieval run was \emph{successful}.
To answer \textbf{RQ2.1}, we register the number configuring I/O examples for each run.
The structural categories of the configuring I/O examples are saved to answer \textbf{RQ2.3}.
Recall and precision of those techniques vary with the number of lines selected for retrieval.
We evaluate the impact of varying the number of extracted lines by multiplying the average number of lines present in the configuring examples with a \emph{retrieval size factor} from 0.5 to 2.5 in steps of 0.5.

@@

\subsection{Common Text Similarity (CTS)}
Figure~\ref{fig:recall-precision-examplecount-TS} presents precision and recall of chunk retrieval using CTS for different numbers of configuring I/O examples.
The count of configuring I/O examples has no visible influence on recall and precision of the chunk retrieval with CTS.

@@

Figure~\ref{fig:contextsizefactor-precision-recall-TS} shows the impact of the retrieval size factor on precision and accuracy of chunk retrieval runs with CTS.
The precision ranges from 52\% when retrieving half expected number of lines to 25\% when 2.5 times the expected number of lines.
The recall ranges from 30\% to 55\%.

@@

Figure~\ref{fig:contextsizefactor-precision-recall-SKWS} shows the impact of the retrieval size factor on precision and accuracy of chunk retrieval runs with SKWS.
The precision ranges from 19\% when retrieving half expected number of lines to 10\% when 2.5 times the expected number of lines.
The recall ranges from 58\% to 75\%.

Figure~\ref{fig:retrievalsizefactor-retrievalratio-SKWS} shows the average proportion of the lines in the test log file retrieved for different retrieval size factors with KWS.
When retrieving the mean number of lines in the configuring I/O examples, 15\% of the log lines are retrieved on average.
This proportion goes up to 25\% when retrieving 1.5 times more lines.

@@

The influence of a single structural category present in the configuring I/O examples compared to multiple categories present is shown in Figure~\ref{fig:recall-precision-singlecategory-all} and Figure~\ref{fig:recall-precision-multicategory-all}.
For more than one present category, precision and recall of PBE greatly decrease.
For CTS and KWS the values also decrease slightly, while RLR is not impacted by the number of structural categories present.

>>>>

The examples are sorted chronologically, i.e.\ a technique is configured with examples from the directly preceding build logs.
For each example set and each technique, we select one to five consecutive I/O examples as configuration, the \emph{configuring I/O examples}, and run the chunk retrieval on the following I/O example, the \emph{test I/O example}.

For each of our evaluation runs we measure the desired and the actual output lines of the chunk retrieval.
The desired output is given by the output of the test I/O example, which acts as the oracle for our evaluation.
From this we obtain true/false positives/negatives and calculate precision, recall and accuracy.
We use these values to answer \textbf{RQ2.2}.
If all desired lines are extracted, i.e.\ recall is one, we say a chunk retrieval run was \emph{successful}.
To answer \textbf{RQ2.1}, we register the number configuring I/O examples for each run.
The structural categories of the configuring I/O examples are saved to answer \textbf{RQ2.3}.
Recall and precision of those techniques vary with the number of lines selected for retrieval.
We evaluate the effect of varying the number of extracted lines by multiplying the average number of lines present in the configuring examples with a \emph{retrieval size factor} from 0.5 to 2.5 in steps of 0.5.

@@

\subsection{Common Text Similarity (CTS)}
Figure~\ref{fig:recall-precision-examplecount-TS} presents precision and recall of chunk retrieval using CTS for different numbers of configuring I/O examples.
The count of configuring I/O examples has no visible influence on recall and precision of the chunk retrieval with CTS\@.

@@

Figure~\ref{fig:contextsizefactor-precision-recall-TS} shows the effect of the retrieval size factor on precision and accuracy of chunk retrieval runs with CTS\@.
The precision ranges from 52\% when retrieving half expected number of lines to 25\% when 2.5 times the expected number of lines.
The recall ranges from 30\% to 55\%.

@@

Figure~\ref{fig:contextsizefactor-precision-recall-SKWS} shows the effect of the retrieval size factor on precision and accuracy of chunk retrieval runs with KWS\@.
The precision ranges from 19\% when retrieving half expected number of lines to 10\% when 2.5 times the expected number of lines.
The recall ranges from 58\% to 75\%.

Figure~\ref{fig:retrievalsizefactor-retrievalratio-SKWS} shows the average proportion of the lines in the test log file retrieved for different retrieval size factors with KWS\@.
When retrieving the mean number of lines in the configuring I/O examples, 15\% of the log lines are retrieved on average.
This proportion goes up to 25\% when retrieving 1.5 times more lines.

@@

Figure~\ref{fig:recall-precision-singlecategory-all} and Figure~\ref{fig:recall-precision-multicategory-all} show the influence of a single structural category present in the configuring I/O examples compared to multiple categories present.
For more than one present category, precision and recall of PBE greatly decrease.
For CTS and KWS the values also decrease slightly, while RLR is not affected by the number of structural categories present.

### 2019-11-06 thesis/txt/6-discussion.tex
adress some linter warnings from lacheck and style-check.rb

When all configuring I/O Examples are of the same structure, one to three examples are enough input for PROSE to synthesize a correct regular expression program. In our study, additional examples did not improve the quality of teh chunk retrieval.

\paragraph{Retrieval Output Usage}
When the program synthesis succeeds and applying the regular expression program yields an output, PBE shows a high precision and high recall for chunk retrieval.
A failure in the program synthesis or no output from the synthesized program is clearly identified by the tool.
Therefore, if there is an output, the user can have high confidence it is the correct output.
This makes output from PBE chunk retrieval well suited for consumption of other software components,

@@

Extracting the average number of lines present in the configuring I/O examples seems to give the best balance of precision and recall for CTS.

@@

\subsection{Keyword Search (KWS)}
\paragraph{Configuration and Input}
The number of structural categories in the configuring I/O examples have little impact on the extraction quality of KWS.
This makes KWS a good technique if there is little prior knowledge on how the desired BLI is represented in the build log to be analyzed.
For the example of extracting the build failure reason, KWS is best suited if a build can fail in various steps logged by different tools and no pre-categorization of where the build failed is available.

@@

\paragraph{Retrieval Output Usage}
Even though KWS has the highest recall of all three techniques, its precision is also the lowest.
The output of a chunk retrieval with KWS is well suited to be read by humans, matching our 
very low precision, only useful to filter lines in build log for further processing, e.g. reading by a developer

\section{Choosing a Suitable Chunk Retrieval Technique}
This section unifies our study results of the chunk retrieval techniques PBE, CTS and KWS.
It recommends which technique to use depending on the available configuration and 
TABLE!
\paragraph{Configuration and Input}
number of categories:
one? -> pbe
two? -> CTS
more? -> KWS
available examples:
little? -> try pbe (This depends on the categories! here we might want a decision tree to represent the importance of these characteristics against each other? <- this would also anwser RQ1: which ones of the criteria we compare are the most important)
multiple -> kws
very many? -> CTS

@@

\begin{table}[htbp]
\begin{tabular}{ |c||c|c|c| }
  Technique & \multicolumn{3}{|c|}{Number of Categories in configuring examples} \\
  \hline
  & one & two & more \\
  PBE & yes & no & no \\ 
  CTS & yes & yes & no \\ 
  KWS & yes & yes & yes \\ 
  \hline
\end{tabular}
\caption{The small $1 \cdot 1$}
\label{tab:101}
\end{table}

>>>>

When all configuring I/O Examples are of the same structure, one to three examples are enough input for PROSE to synthesize a correct regular expression program.
In our study, additional examples did not improve the quality of the chunk retrieval.

\paragraph{Retrieval Output Usage}
When the program synthesis succeeds and applying the regular expression program yields an output, PBE shows a high precision and high recall for chunk retrieval.
A failure in the program synthesis or no output from the synthesized program is clearly identified by the tool.
Therefore, if there is an output, the user can have high confidence it is the correct output.
This makes output from PBE chunk retrieval well suited for consumption of other software components.

@@

Extracting the average number of lines present in the configuring I/O examples seems to give the best balance of precision and recall for CTS\@.

@@

\subsection{Keyword Search (KWS)}
\paragraph{Configuration and Input}
The number of structural categories in the configuring I/O examples have little impact on the extraction quality of KWS\@.
This makes KWS a good technique if there is little prior knowledge on how the desired BLI is represented in the build log to be analyzed.
For the example of extracting the build failure reason, KWS is best suited if a build can fail in various steps logged by different tools and no pre-categorization of where the build failed is available.

@@

\paragraph{Retrieval Output Usage}
Even though KWS has the highest recall of all three techniques, its precision is also the lowest.
The output of a chunk retrieval with KWS is well suited to be read by humans.

\section{Choosing a Suitable Chunk Retrieval Technique}
This section unifies our study results of the chunk retrieval techniques PBE, CTS and KWS\@.
It recommends which technique to use depending on the available configuration and 
TABLE!
\paragraph{Configuration and Input}
number of categories:
one? -> pbe
two? -> CTS
more? -> KWS
available examples:
little? -> try pbe (This depends on the categories! here we might want a decision tree to represent the importance of these characteristics against each other? <- this would also anwser RQ1: which ones of the criteria we compare are the most important)
multiple -> kws
very many? -> CTS

@@

\begin{table}[htbp]
\begin{tabular}{|c||c|c|c|}
  Technique & \multicolumn{3}{|c|}{Number of Categories in configuring examples} \\
  \hline
  & one & two & more \\
  PBE & yes & no & no \\ 
  CTS & yes & yes & no \\ 
  KWS & yes & yes & yes \\ 
  \hline
\end{tabular}
\caption{The small $1 \cdot 1$}
\label{tab:101}
\end{table}
